---
title: "665final"
author: "Ty Darnell"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=F,warning=F)
```

```{r,echo=F}
library(knitr)
library(survival)
library(tidyverse)
```

# Part 1

# 1

Without using a formal statistical model, provide an estimate of the common odds ratio and its 95%
confidence interval for the effect of pooled treatment (high dose + low dose) vs. placebo on the
severity of the adverse event, dichotomized as (none or mild) vs. (moderate or severe), when
controlling for sex.

The Common odds ratio is 2.146 with a 95% CI $(1.370,3.361)$ 

Subjects with in the pooled treatment group have 2.146 times the odds of having a moderate or severe adverse event (vs none or mild) compared to the odds of a subject in the placebo group, when controlling for sex.

# 2

Without using a formal statistical model, statistically test the null hypothesis that the effect of pooled
treatment (high dose + low dose) vs. placebo on the dichotomized severity of the adverse event ̶
(none or mild) vs. (moderate or severe) ̶ is the same for each sex. Provide a sentence explaining
your results.


# 3 

Under minimal assumptions, conduct a statistical test to determine whether there is a difference in
the proportion of moderate or severe adverse event (vs. none or mild) among the three treatment
groups, controlling for sex. For this problem, you should consider the treatment groups as nominal.
Write a sentence to interpret your findings.

# 4

Under minimal assumptions, conduct a statistical test to determine whether there is a trend in the
proportion of moderate or severe adverse event (vs. none or mild) across the ordered treatment
groups, controlling for sex. Write a sentence to interpret your findings.

# 5

Under minimal assumptions, conduct a statistical test to assess the association of pooled treatment
(high dose + low dose) vs. placebo with the severity of adverse event (all four ordered levels managed
as distinct), controlling for sex. Justify your method. If you determine that p < 0.05, discuss whether
pooled treatment is associated with greater severity or lesser severity of the adverse event.

# 6

Under minimal assumptions, conduct a statistical test to determine whether there is a progressive
location shift in the severity of the adverse event (as distinct levels) across high dose, low dose, and
placebo, controlling for sex. In a sentence, interpret your findings.


# 7

Report the Spearman rank correlation coefficients and corresponding 95% confidence intervals
separately by sex as measures of association for pooled treatment (high dose + low dose) versus
placebo with the severity of adverse event (as distinct levels). Write a sentence indicating whether
men or women exhibit a stronger association, and briefly justify your finding.

# 8

Report the Spearman rank correlation coefficients and corresponding 95% confidence intervals
separately by sex as measures of association for ordered treatment groups with severity of the adverse
event (as distinct levels). Write a sentence indicating whether men or women exhibit a stronger
association, and briefly justify your finding.

# 9 

Separately within each treatment group, test the association between sex and severity of the adverse
event (as ordered distinct levels). Also, assess such association under minimal assumptions, and
controlling for treatment groups. Write a sentence to interpret your results.

# 10

Mathematically specify a logistic regression model for the dichotomous response of (moderate or
severe) adverse event (vs. none or mild), with main effects for treatment (as distinct levels and
considered as nominal, treating placebo as the reference) and sex (treating men as the reference).
State assumptions, and mathematically define all variables in the model. Interpret all model
parameters. Determine the goodness of fit for this model.

# 11

Using the model from Problem 10, provide estimates and corresponding 95% confidence intervals
for the odds ratios of high dose vs. placebo and of low dose vs. placebo for (moderate or severe)
adverse event compared to (none or mild).

# 12

Using the model from Problem 10, perform a statistical test of whether the treatment groups differ
with respect to the (moderate or severe) adverse event (i.e., the overall treatment effect). Provide the
test statistic, indicate the number of degrees of freedom, and determine statistical significance
through its p-value. If this overall effect is statistically significant, test each pairwise treatment
comparison at the α=0.05 level, and indicate which treatment groups are significantly better than
others. (Note: you do not need to address any adjustment to the type I error for multiple comparisons
for this problem.)

# 13

Using your model from Problem 10, what are the respective model-predicted probabilities for
(moderate or severe) adverse event and for (none or mild) adverse event for men on high dose and
also for women on placebo?

# 14

Mathematically specify a proportional odds regression model for ‘more’ vs. ‘less’ severity of the
adverse event, with main effects for treatment (as distinct levels and considered as nominal, treating
placebo as the reference) and sex (treating men as the reference). State assumptions, and
mathematically define all variables in the model. Interpret all model parameters. Determine the
goodness of fit for this model, including the test for proportional odds. If the proportional odds
assumption is not supported, investigate and briefly describe your findings regarding the potential
reason for this.

# 15 

Using the model from Problem 14 that assumes proportional odds, provide estimates and
corresponding 95% confidence intervals for the odds ratios of high dose vs. placebo and of low dose
vs. placebo for (severe or moderate) adverse event compared to (none or mild).

# 16

Using the model from Problem 14 that assumes proportional odds, what are the respective model-
predicted probabilities for none, mild, moderate, and severe adverse event for men on low dose?

# 17

Mathematically specify a generalized logit model for all levels of severity of adverse event (treating
‘none’ as the reference), with main effects for treatment (as distinct levels and considered as nominal,
treating placebo as the reference) and sex (treating men as the reference). State assumptions, and
mathematically define all variables in the model. Interpret all model parameters.

# 18

Using your model from Problem 17, provide estimates and corresponding 95% confidence intervals
for the odds ratios of high dose vs. placebo and of low dose vs. placebo for each severity level
compared to ‘none’. Compare and contrast these results to those found in Problem 15.

# Part 2
